Effects of Reynolds Number on the Energy Conversion and Near-Wake Dynamics of a High Solidity Vertical-Axis Cross-Flow Turbine

### 2016-01-25 paper.tex
Modifications from editor, round 2

% Contact information of the corresponding author (Add [2] after \corres if
% there are more than one corresponding author.)
\corres[2]{\hspace{-1em} Correspondence: pxL3@unh.edu (P.B.);
    martin.wosnik@unh.edu (M.W.); \linebreak Tel.: +1-603-862-1891 (M.W.)}

@@

% Keywords: add 3 to 10 keywords
\keyword{\textls[-10]{Reynolds number; cross-flow turbine; turbine performance; marine
hydrokinetic energy}; wind energy; {VAWT}; scale model}

@@

Scaled physical models are often used in science and engineering to approximate
real-world systems. The principle of dynamic similarity allows for
geometrically-scaled systems to be dynamically identical if certain
nondimensional physical parameters are matched. In the case of fluid systems,
the most important nondimensional parameter is often the Reynolds number, $Re$,
which quantifies the importance of inertial forces over viscous forces on the
flow physics \cite{Acheson1990}: $Re = Ul/\nu$, where $U$ and $l$ are
characteristic velocity and length scales, respectively, and $\nu$ is the fluid
kinematic viscosity. The~other common dynamical scale, for systems with a free
surface, is the Froude number $Fr = U/\sqrt{gl}$, where $g$ is the gravitational
acceleration. Matching both the Froude and Reynolds number is not possible for a
scaled model in a given fluid, which is illustrated by the relation $Re =
l^{3/2} g^{1/2} Fr / \nu$, since $Re$ scales linearly with $l$, or the geometric
scale. In general, the geometric scale of a physical model is not necessarily
the same as its dynamical scale. As such, hereafter, we will use the word
``scale'' to refer to this dynamical scale or the Reynolds number and assume the
Froude number is small enough to be negligible.

@@

Blackwell \emph{et al.} \cite{Blackwell1976} investigated the effects of
Reynolds number on the performance of a 2 m diameter Darrieus vertical-axis
cross-flow wind turbine with NACA 0012 blades. By varying wind tunnel speed, the
turbine was made to operate at an approximately constant blade chord Reynolds
number $Re_c$ ranging from $1 \times 10^5$ to $3 \times 10^5$. In this regime,
the turbine power coefficient $C_P$ was shown to be sensitive to $Re_c$, with
sensitivity diminishing at the higher Reynolds numbers, especially for turbines
of lower solidity ($Nc/R$, where $N$ is the number of blades and $R$ is the
turbine's maximum radius). More recently, Polagye and Cavagnaro
\cite{Polagye2013b} observed significant Reynolds number sensitivity for a high
solidity helical cross-flow turbine, and Bravo \emph{et al.} \cite{Bravo2007}
observed the power coefficient of a straight-bladed VAWT to become Reynolds
number independent at $Re_c \approx 4 \times 10^5$. Bachant and Wosnik
\cite{Bachant2014} observed Reynolds number independence of the power
coefficient at the optimal tip speed ratio for a high solidity cross-flow
turbine at $Re_D \approx 10^6$ or $Re_{c,\mathrm{avg}} \approx 2 \times 10^5$.

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=0.7\textwidth]{figures/exp-setup-photo}
        \caption{}
        \label{fig:exp-setup-photo}
    \end{subfigure}

    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=0.7\textwidth]{figures/exp_setup_drawing}
        \caption{}
        \label{fig:exp-setup-dwg}
    \end{subfigure}

    \caption{Experimental setup photo (a) and drawing (b): turbine test bed
        installed in the UNH tow tank.}

@@

\includegraphics[width=0.9\textwidth]{figures/turbine_coordinate_system}

@@

\includegraphics[width=0.9\textwidth]{figures/perf_re_dep}

\caption{UNH-RVAT measured mean power \textbf{(a)} and drag \textbf{(b)}
    coefficients at $\lambda=1.9$ plotted \textit{versus} the Reynolds number. Error
    bars indicate expanded uncertainty estimates for 95\% confidence, which for
    $C_P$ is dominated by systematic error estimates from the torque transducer
    operating at the lower end of its measurement range.}

@@

To help understand, and possibly predict, the $Re$-sensitivity on turbine
performance, a series of static foil coefficient datasets were computed with the
viscous panel method code XFOIL \cite{Drela1989}, a commonly-used tool for
airfoil analysis, e.g., \cite{Castelli2011, Walker2014}, implemented as part of
the open source turbine design software QBlade \cite{Marten2013}. Simulations
were run for an angle of attack range of 0$^{\circ}$ to 40$^{\circ}$, in
increments of 0.5$^{\circ}$. Solver parameters used were 100 panels, a fixed
speed, zero Mach number, \texttt{NCrit = 9} (default $\mathrm{e}^n$ transition
criteria parameter for an average wind tunnel) and no forced boundary layer
transition. Characteristics were computed for the approximate average blade
chord Reynolds numbers encountered in the tow tank experiments.

@@

    \includegraphics[width=0.9\textwidth]{figures/all_foils_re_dep}

    \caption{Normalized maximum lift coefficient \textbf{(a)}, drag coefficient
        \textbf{(b)}, and lift-to-drag ratio \textbf{(c)} computed by XFOIL at
        various $Re_c$  for each profile.}

@@

    \caption{Geometric angle of attack \textbf{(a)}, relative velocity
        \textbf{(b)},  and torque coefficient \textbf{(c)} calculated with a NACA
        0020 foil operating at $\lambda=1.9$.}

@@

    \includegraphics[width=0.8\textwidth]{figures/cft_re_dep_foils}

    \caption{Reynolds number dependence of the normalized peak torque
        coefficient calculated from static foil coefficients and blade kinematics,
        compared to experimental data from a cross-flow turbine. Note that the
        experimental data represents the mean torque coefficient, not the maximum.}

@@

     \caption{Mean streamwise velocity \textbf{(a)} and turbulence kinetic
         energy \textbf{(b)} profiles at $z/H=0.0$. Turbine diameter Reynolds number
         $Re_D$ is indicated by the legend.}

@@

\vspace{6pt} \acknowledgments{\textbf{Acknowledgments:} The authors would like
    to acknowledge funding through a National Science Foundation CAREER award
    (principal investigator Martin Wosnik, NSF 1150797, Energy for Sustainability,
    program manager Gregory L. Rorrer), a grant through the Leslie S. Hubbard Marine
    Program Endowment to purchase acoustic flow measurement instrumentation, and a
    grant for laboratory infrastructure upgrades through the U.S. Department of
    Energy.}

>>>>

% Contact information of the corresponding author (Add [2] after \corres if
% there are more than one corresponding author.)
\corres[2]{\hspace{-1em} Correspondence: pxL3@unh.edu (P.B.);
    martin.wosnik@unh.edu (M.W.); \linebreak Tel.: +1-603-862-1891 \hl {(M.W.)}}
    %please also provide the author (P.B.)'s office number. and please also pay attention to the keywords section, can we define VAWT as vertical-axis wind turbine?

@@

% Keywords: add 3 to 10 keywords
\keyword{\textls[-10]{Reynolds number; cross-flow turbine; turbine performance; marine
hydrokinetic energy}; wind energy; {vertical-axis wind turbine (VAWT)}; scale model}

@@

Scaled physical models are often used in science and engineering to approximate
real-world systems. The principle of dynamic similarity allows for
geometrically-scaled systems to be dynamically identical if certain
nondimensional physical parameters are matched. In the case of fluid systems,
the most important nondimensional parameter is often the Reynolds number, $Re$,
which quantifies the importance of inertial forces over viscous forces on the
flow physics \cite{Acheson1990}: $Re = Ul/\nu$, where $U$ and $l$ are
characteristic velocity and length scales, respectively, and $\nu$ is the fluid
kinematic viscosity. The~other common dynamical scale, for systems with a free
surface, is the Froude number $Fr = U/\sqrt{gl}$, where $g$ is the gravitational
acceleration. Matching both the Froude and Reynolds number is not possible for a
scaled model in a given fluid, which is illustrated by the relation $Re =
l^{3/2} g^{1/2} Fr / \nu$, since $Re$ scales linearly with $l$, or the geometric
scale. In general, the geometric scale of a physical model is not necessarily
the same as its dynamical scale. As such, hereafter, we will use the word
``scale'' to refer to this dynamical scale or the Reynolds number and assume the
Froude number is small enough to be~ negligible.

@@

Blackwell \emph{et al.} \cite{Blackwell1976} investigated the effects of
Reynolds number on the performance of a 2 m diameter Darrieus vertical-axis
cross-flow wind turbine with NACA 0012 blades. By varying wind tunnel speed, the
turbine was made to operate at an approximately constant blade chord Reynolds
number $Re_c$ ranging from $1 \times 10^5$ to $3 \times 10^5$. In this regime,
the turbine power coefficient $C_P$ was shown to be sensitive to $Re_c$, with
sensitivity diminishing at the higher Reynolds numbers, especially for turbines
of lower solidity ($Nc/R$, where $N$ is the number of blades and $R$ is the
turbine's maximum radius). More recently, Polagye and Cavagnaro
\cite{Polagye2013b} observed significant Reynolds number sensitivity for a high
solidity helical cross-flow turbine, and Bravo \emph{et al.} \cite{Bravo2007}
observed the power coefficient of a straight-bladed \hl {vertical-axis wind turbine (VAWT)} to become Reynolds
number independent at $Re_c \approx 4 \times 10^5$. Bachant and Wosnik
\cite{Bachant2014} observed Reynolds number independence of the power
coefficient at the optimal tip speed ratio for a high solidity cross-flow
turbine at $Re_D \approx 10^6$ or $Re_{c,\mathrm{avg}} \approx 2 \times 10^5$.

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=0.6\textwidth]{figures/exp-setup-photo}
        \caption{}
        \label{fig:exp-setup-photo}
    \end{subfigure}

    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=0.6\textwidth]{figures/exp_setup_drawing}
        \caption{}
        \label{fig:exp-setup-dwg}
    \end{subfigure}

    \caption{Experimental setup photo (a) and drawing (b): turbine test bed
        installed in the UNH tow~ tank.}

@@

\includegraphics[width=0.8\textwidth]{figures/turbine_coordinate_system}

@@

\includegraphics[width=0.9\textwidth]{figures/perf_re_dep}
%Please use scientific notation in the figure. for instance, please revise 1e6 to \times 10^6
\caption{\hl {UNH-RVAT} measured mean power (\textbf{a}) and drag (\textbf{b})
    coefficients at $\lambda=1.9$ plotted \textit{versus} the Reynolds number. Error
    bars indicate expanded uncertainty estimates for 95\% confidence, which for
    $C_P$ is dominated by systematic error estimates from the torque transducer
    operating at the lower end of its measurement range.}

@@

To help understand, and possibly predict, the $Re$-sensitivity on turbine
performance, a series of static foil coefficient datasets were computed with the
viscous panel method code XFOIL \cite{Drela1989}, a commonly-used tool for
airfoil analysis, e.g., \cite{Castelli2011, Walker2014}, implemented as part of
the open source turbine design software QBlade \cite{Marten2013}. Simulations
were run for an angle of attack range of 0$^{\circ}$ to 40$^{\circ}$, in
increments of 0.5$^{\circ}$. Solver parameters used were 100 panels, a fixed
speed, zero Mach number, \linebreak \texttt{NCrit = 9} (default $\mathrm{e}^n$ transition
criteria parameter for an average wind tunnel) and no forced boundary layer
transition. Characteristics were computed for the approximate average blade
chord Reynolds numbers encountered in the tow tank experiments.

@@

    \includegraphics[width=0.9\textwidth]{figures/all_foils_re_dep}
 % Please use scientific notation in the figure. for instance, please revise 1e5 to \times 10^5 
    \caption{\hl {Normalized maximum} lift coefficient (\textbf{a}), drag coefficient
        (\textbf{b}), and lift-to-drag ratio (\textbf{c}) computed by XFOIL at
        various $Re_c$  for each profile.}

@@

    \caption{Geometric angle of attack (\textbf{a}), relative velocity
        (\textbf{b}),  and torque coefficient (\textbf{c}) calculated with a NACA
        0020 foil operating at $\lambda=1.9$.}

@@

    \includegraphics[width=0.8\textwidth]{figures/cft_re_dep_foils}
%Please use scientific notation in the figure. for instance, please revise 1e6 to \times 10^6
    \caption{\hl {Reynolds} number dependence of the normalized peak torque
        coefficient calculated from static foil coefficients and blade kinematics,
        compared to experimental data from a cross-flow turbine. Note that the
        experimental data represents the mean torque coefficient, not the maximum.}

@@

     \caption{Mean streamwise velocity (\textbf{a}) and turbulence kinetic
         energy (\textbf{b}) profiles at $z/H=0.0$. Turbine diameter Reynolds number
         $Re_D$ is indicated by the legend.}

@@

\vspace{6pt} \acknowledgments{\textbf{Acknowledgments:} The authors would like
    to acknowledge funding through a National Science Foundation CAREER award
    (principal investigator Martin Wosnik, NSF 1150797, Energy for Sustainability,
    program manager Gregory L. Rorrer), a grant through the Leslie S. Hubbard Marine
    Program Endowment to purchase acoustic flow measurement instrumentation, and a
    grant for laboratory infrastructure upgrades through the U.S. Department of~
    Energy.}
